Methods for Analyzing Models of Rod-Shaped Bacteria

### 2025-07-26 paper/main.tex
insert figures in place
remove bullet points
introduction to model

\begin{enumerate}
    \item represent bacteria as polygon with vertices $\vec{v}_i$
    \item Assumes that these vertices are not fundamental but rather a discretization of the actual
        problem
\end{enumerate}

@@

\begin{figure}
    \centering
    \begin{tikzonimage}[width=0.36\textwidth]
        {docs/source/_static/fitting-methods/algorithm/image001042.png}
        \node at (0.025, 0.975)[anchor=north west, rectangle, draw, white, minimum width=15pt, minimum height=15pt]{\textbf{A}};
    \end{tikzonimage}
    \begin{tikzonimage}[width=0.36\textwidth]
        {docs/source/_static/fitting-methods/algorithm/mask-zoom.png}
        \node at (0.025, 0.975)[anchor=north west, rectangle, draw, white, minimum width=15pt, minimum height=15pt]{\textbf{B}};
    \end{tikzonimage}
    \begin{tikzonimage}[width=0.255\textwidth]
        {docs/source/_static/fitting-methods/algorithm/interpolate-positions.png}
        \node at (0.025, 0.975)[anchor=north west, rectangle, draw, white, minimum width=15pt, minimum height=15pt]{\textbf{C}};
    \end{tikzonimage}
    \caption{
        (A) Input image.
        (B) Result of the skeletonization algorithm for a single cell.
        (C) Vertices $\vec{v}_i$ which have been interpolated from the skeletonized cell.
        They represent the position of our cell-agents.
    }
    \label{fig:position-extraction-algorithm}
\end{figure}

@@

\begin{figure}
    \centering
    \begin{tikzonimage}[width=0.32\textwidth]
        {docs/source/_static/fitting-methods/extract_positions-004000.png}
        \node at (0.025, 0.975)[anchor=north west, rectangle, draw, white, minimum width=15pt, minimum height=15pt]{\textbf{A}};
    \end{tikzonimage}
    \begin{tikzonimage}[width=0.32\textwidth]
        {docs/source/_static/fitting-methods/extract_positions-006000.png}
        \node at (0.025, 0.975)[anchor=north west, rectangle, draw, white, minimum width=15pt, minimum height=15pt]{\textbf{B}};
    \end{tikzonimage}
    \begin{tikzonimage}[width=0.32\textwidth]
        {docs/source/_static/fitting-methods/extract_positions-008000.png}
        \node at (0.025, 0.975)[anchor=north west, rectangle, draw, white, minimum width=15pt, minimum height=15pt]{\textbf{C}};
    \end{tikzonimage}
    \begin{tikzonimage}[width=0.5\textwidth]
        {docs/source/_static/fitting-methods/displacement-calculations.png}%
        \node at (0.03, 0.99)[anchor=north west, rectangle, draw, black, minimum width=15pt, minimum height=15pt]{\textbf{D}};
    \end{tikzonimage}%
    \begin{tikzonimage}[width=0.5\textwidth]
        {docs/source/_static/fitting-methods/displacement-distribution.png}
        \node at (0.03, 0.99)[anchor=north west, rectangle, draw, black, minimum width=15pt, minimum height=15pt]{\textbf{E}};
    \end{tikzonimage}
    \caption{
        Progression of generated masks with estimated and exact positions.
        Comparison of masks across generations. Compare estimated with exact rod lengths.
"Iceberg plot"
    }
    \label{fig:benchmarking-extraction-algorithm}
\end{figure}

@@

%###################################################################################################

In order to construct a mechanical model of rod-shaped bacteria, we have to take care of a number of
different aspects.
The mechanical behaviour of rod-shaped bacteria is governed by a plethora of effects.
Although we set out to construct a model which is able to describe as many of these effects as
possible, we will fall short in this paper and restrict ourselves to the aspects presented
in~\ref{table:simulation-aspects}.
We will discuss, how this model can be extended in order to capture more effects and how we can
validate our approach given the existing literaturesection:supplement-more-simulation-aspects
(see section~\ref{section:discussion}, and~\ref{section:supplement-more-simulation-aspects}).

>>>>

% \begin{enumerate}
%     \item represent bacteria as polygon with vertices $\vec{v}_i$
%     \item Assumes that these vertices are not fundamental but rather a discretization of the actual
%         problem
% \end{enumerate}

@@

\begin{figure}[H]
    \centering
    \begin{tikzonimage}[width=0.36\textwidth]
        {docs/source/_static/fitting-methods/algorithm/image001042.png}
        \node at (0.025, 0.975)[anchor=north west, rectangle, draw, white, minimum width=15pt, minimum height=15pt]{\textbf{A}};
    \end{tikzonimage}
    \begin{tikzonimage}[width=0.36\textwidth]
        {docs/source/_static/fitting-methods/algorithm/mask-zoom.png}
        \node at (0.025, 0.975)[anchor=north west, rectangle, draw, white, minimum width=15pt, minimum height=15pt]{\textbf{B}};
    \end{tikzonimage}
    \begin{tikzonimage}[width=0.255\textwidth]
        {docs/source/_static/fitting-methods/algorithm/interpolate-positions.png}
        \node at (0.025, 0.975)[anchor=north west, rectangle, draw, white, minimum width=15pt, minimum height=15pt]{\textbf{C}};
    \end{tikzonimage}
    \caption{
        (A) Input image.
        (B) Result of the skeletonization algorithm for a single cell.
        (C) Vertices $\vec{v}_i$ which have been interpolated from the skeletonized cell.
        They represent the position of our cell-agents.
    }
    \label{fig:position-extraction-algorithm}
\end{figure}

@@

\begin{figure}[H]
    \centering
    \begin{tikzonimage}[width=0.32\textwidth]
        {docs/source/_static/fitting-methods/extract_positions-004000.png}
        \node at (0.025, 0.975)[anchor=north west, rectangle, draw, white, minimum width=15pt, minimum height=15pt]{\textbf{A}};
    \end{tikzonimage}
    \begin{tikzonimage}[width=0.32\textwidth]
        {docs/source/_static/fitting-methods/extract_positions-006000.png}
        \node at (0.025, 0.975)[anchor=north west, rectangle, draw, white, minimum width=15pt, minimum height=15pt]{\textbf{B}};
    \end{tikzonimage}
    \begin{tikzonimage}[width=0.32\textwidth]
        {docs/source/_static/fitting-methods/extract_positions-008000.png}
        \node at (0.025, 0.975)[anchor=north west, rectangle, draw, white, minimum width=15pt, minimum height=15pt]{\textbf{C}};
    \end{tikzonimage}
    \begin{tikzonimage}[width=0.5\textwidth]
        {docs/source/_static/fitting-methods/displacement-calculations.png}%
        \node at (0.03, 0.99)[anchor=north west, rectangle, draw, black, minimum width=15pt, minimum height=15pt]{\textbf{D}};
    \end{tikzonimage}%
    \begin{tikzonimage}[width=0.5\textwidth]
        {docs/source/_static/fitting-methods/displacement-distribution.png}
        \node at (0.03, 0.99)[anchor=north west, rectangle, draw, black, minimum width=15pt, minimum height=15pt]{\textbf{E}};
    \end{tikzonimage}
    \caption{
        Progression of generated masks with estimated and exact positions.
        Comparison of masks across generations. Compare estimated with exact rod lengths.
"Iceberg plot"
    }
    \label{fig:benchmarking-extraction-algorithm}
\end{figure}

@@

The mechanical behaviour of rod-shaped bacteria is governed by a plethora of effects.
In this work, we aim to establish a model which is abstract enough such that it can be reused for
many types of rod-shaped bacteria but also explicit enough such that we can quantify its parameters.
We are thus aiming to work with a set of assumptions that is as general as possible but still
specific to rod-shaped bacteria.
With this delicate balance in mind, it is clear that many cellular processes which might be specific
to one particular celltype or simply not necessary to assume in order to construct our model, will
be omitted alltogether.
This concerns effects such as polar interactions~\cite{Duvernoy2018}, extracellular
reactions~\cite{Wang2010} or differentiation processes~\cite{vanGestel2015,Lpez2010} and many other.
We restrict ourselves to the aspects presented in table~\ref{table:simulation-aspects}.
For a collection of other known effects which could be considered, we refer to the supplementary
section~\ref{section:discussion}, and~\ref{section:supplement-more-simulation-aspects}.
